---
title: "A Multinational Bank Customer Churn"
author: "T4-Analytical Avengers"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
```{r init, include=FALSE}
library(ezids)
loadPkg("Hmisc") 
library(dplyr)
loadPkg("MASS")
library(ggplot2)
library(pkgload)
library(corrplot)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
```

# EDA

**Reading the file with all the customer churn details as a data frame.Then displaying the structure of the data frame**

```{r Churn_data_file, include=T,echo=TRUE,results="rpois"}
churn_data<-data.frame(read.csv("Bank_Customer_Churn_Prediction.csv",header=TRUE)) #reading the file into data frame
str(churn_data) #printing the data frame structure
```

The churn data frame as `r nrow(churn_data)` customers from the  with  each customer  having `r ncol(churn_data)`  following attributes;
Customer ID :The Unique ID of each individual customer

Credit Score : A number depicting the customer's creditworthiness

Country : The country the customer banks from

Gender : The gender the customer identifies with

Age : Depicts the customers age

Tenure : Indicates how length in years the customer has been with the bank

Balance : The amount currently available in the customer's account

Products Number : The number of products purchased by the customer through the bank

Credit Card : Indicates the customer has a credit card

Active Member : Indicates if the customer is an active or inactive

Estimated Salary : Bank Estimation of the income of the customer

Churn : Indicator of if the customer has left the bank or not


**Checking for duplicates in the data frame.**
```{r duplicates}
nrow(unique(churn_data)) #checking for unique rows
```

There are `r nrow(unique(churn_data))` rows of unique records in the data.Hence we the data frame has no duplicates records.We can proceed with next step of analysis.

**Checking if there are any NA values in the data frame**
```{r Check for NA in the dataframe}
sum(is.na(churn_data) == 'TRUE') # checking for NA values in the dataframe

```
We found that our dataset was clean without any NA values.

**Checking for summary of the churn data frame.**

```{r churn_summary,results='apois'}
summary(churn_data) #looking for summary of the data frame
```
Credit score:We see that minimum credit score of the customer is 350 with average score being 651 and 850 being maximum score.
Age:Minimum age of the customer is 18 with average age being 38.9 and maximum age being 92.
Tenure:The customers has been with bank for an average of 5 years and upto 10 years.
Balance:Maximum account balance of the customer is 250898 with average balance being 76486.
Country,gender,active member,credit card,product n umber and churn are categorical variables.


**Customer ID is not an important attribute that impacts churn rate**
```{r Remove cutomer_id from the dataframe, include=TRUE, echo=TRUE}
churn_data <- churn_data[-c(1)] # removing column 1(customer _id) from data set
ncol(churn_data) # counting number of variables in the data frame
str(churn_data) # checking for summary of the data frame after removing the customer_if field
```
We are not making analysis for each customer,Hence customer_id was irrelevant field in the data set so we decided to remove the customer_id column from the data frame.
After removing customer_id attribute from the data frame we now have `r ncol(churn_data)`  columns.

## Independent variable EDA

**Converting required variables into categorical values**
```{r category}
str(churn_data)
churn_data_category=churn_data
churn_data_category$credit_card=factor(churn_data$credit_card)
churn_data_category$active_member=factor(churn_data$active_member)
churn_data_category$churn=factor(churn_data$churn)
churn_data_category$gender=factor(churn_data$gender)
churn_data_category$tenure =factor(churn_data$tenure )
churn_data_category$products_number=factor(churn_data$products_number)
churn_data_category$age=factor(churn_data$age)
str(churn_data_category)
```

We decided to convert these variables into categorical (factor):  
`credit_card`, `active_member`, `churn` and `gender`,`product_number`,`tenure`,`age`. Notice that 
the data frame `churn_data` still has all variables numerical, while the data frame `churn_data_category` 
include categorical columns that we just converted. 


**Converting Boolean values into Character format**
```{r}
churn_data$active_member[churn_data$active_member== 1]<-"Active"
churn_data$active_member[churn_data$active_member== 0]<-"In Active"

churn_data$credit_card[churn_data$credit_card== 1]<-"Credit Card"
churn_data$credit_card[churn_data$credit_card== 0]<-"No-Credit Card"

churn_data$churn[churn_data$churn== 1]<-"Churned"
churn_data$churn[churn_data$churn== 0]<-"Retained"

head(churn_data)
```
We decided to assign string values to the  corresponding Boolean values for the few variables in the dataset,So that it would be much better to understand each variable clearly.

### Ouliers For Customer data

***Checking for outliers in the data frame  for credit_score,age,tenure,balance,estimated_salary variables  using outlierKD function.***

```{r customer_credit_score}
credit_score_clean <- outlierKD2(churn_data, credit_score, rm = TRUE, boxplt = TRUE, qqplt = TRUE)
xkablesummary(credit_score_clean)
```

The proportion outliers for credit_score variable if 0.2% and it was removed successfully using outlierKD function.

```{r customer_age}
age_clean <- outlierKD2(churn_data, age, rm = TRUE, boxplt = TRUE, qqplt = TRUE)
xkablesummary(age_clean)
```

```{r Cutomer_tenure}
tenure_clean <- outlierKD2(churn_data, tenure, rm = TRUE, boxplt = TRUE, qqplt = TRUE)
xkablesummary(tenure_clean)
```

```{r balance}
balance_clean <- outlierKD2(churn_data, balance, rm = TRUE, boxplt = TRUE, qqplt = TRUE)
xkablesummary(balance_clean)
```

```{r estimated_salary}
salary_clean <- outlierKD2(churn_data, estimated_salary, rm = TRUE, boxplt = TRUE, qqplt = TRUE)
xkablesummary(salary_clean)
```
By using outlierKD function we can observe that outliers where found only in `age` and `credit_score` variables i.e,3.7% and 0.2%,So we decided to remove using outlierKD function.

### Customer Analysis 

Analyzing each variable in the dataframe by using the relevant plots along with the calculated mean,SD and percentages.

```{r Credit_score}

library(ggplot2)
ggplot(data=churn_data, aes(x=credit_score)) + 
  geom_histogram(col="black",bins=30,
                 fill="dark orange", 
                  alpha =.7) + # opacity
  labs(title="`ggplot`") +
  labs(title="Customer's Credit Scores",x="Customer Credit Score", y="Customer Count") 
print(mean(churn_data$credit_score))
print(sd(churn_data$credit_score))

```

**The average `credit score` of the customer is `r mean(churn_data$credit_score)`,most of the customer having credit score fall between 600 to 700 and standard deviation is `r sd(churn_data$credit_score)`.**


```{r Country}
ggplot(data=churn_data,aes(x=country,fill=country))+
  geom_bar(col="black")+
  scale_fill_brewer(palette="Reds") +
  labs(title = "Bank Customer vs Country",x="Country",y="Customer Count")+
  theme_minimal()

```

**The customers are grouped by the countries in which they have their accounts.As we see from the plot France has more than 50% of customer account which is the highest among all other countries with Germany and Spain sharing equal percentages.**


```{r Age}
ggplot(churn_data, aes(x = age,fill=cut(age,100))) + 
  geom_histogram(show.legend = FALSE,col="black",bins=30)+
  scale_fill_discrete(h=c(240,10),c=120,l=70)+
  theme_minimal()+
  labs(x=" Customer Age",y=" Customer Count")+
  ggtitle("Customer ages")
print(mean(churn_data$age))
print(sd(churn_data$age))
```

**The majority of the bank customer's fall below the age of 50 with average age of `r mean(churn_data$age)` and with standard deviation of `r sd(churn_data$age)`.**

```{r tenure}

library(ggplot2)
ggplot(data=churn_data, aes(x=tenure)) + 
  geom_histogram(col="black",bins=11,
                 fill="Yellow", 
                  alpha =.7) + 
  labs(title="Years spent as customer",x="Number of years", y="Customer Count") 

```

**Most of the customer as been with bank for more than a year.**

```{r Gender}
df <- churn_data %>% 
  group_by(gender) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))
str(df)

ggplot(df, aes(x = "", y = perc, fill =gender )) +
  geom_col(color="black") +
  geom_text(aes(label = labels),color = c("black", "black"),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y")+
  labs(title="Percentage of Male and Female customers in bank",x="", y="") 
```  

**The Male customers are highest in percent of 55% and remaining 45% are Female customers.**

```{r Active_Memebers}

df <- churn_data %>% 
  group_by(active_member) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))
str(df)


ggplot(df, aes(x = "", y = perc, fill = active_member)) +
  geom_col(color="black") +
  geom_label(aes(label = labels),color = c("black", "white"),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y")+
  scale_fill_grey()+
  labs(title="Percentage of active members in bank",x="", y="") 
```  

**Huge percentage of customer are being inactive,it is reported that 48.5% of customers are being inactive.**

```{r Credit_Card}
df <- churn_data %>% 
  group_by(credit_card) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))
str(df)

ggplot(df, aes(x = "", y = perc, fill = credit_card)) +
  geom_col(color="black") +
  geom_label(aes(label = labels),color = c("black", "black"),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y")+
  scale_fill_brewer(palette="Greens")+
  labs(title="Percentage of customer use credit card",x="", y="") 
```  

**Predominantly 71% of the Bank customer use credit_card and only 29% do not make use of it.**

```{r Products}
df <- churn_data %>% 
  group_by(products_number) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))
str(df)

ggplot(df, aes(x = "", y = perc, fill = products_number)) +
  geom_col(color="black") +
  geom_label(aes(label = labels),color = c("white", "white","white","white"),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y")+
  labs(title="Percentage of  different products used by customers ",x="", y="") 
  
``` 

**Most of the customer user product 1 which is 50% and product 4 is the least used with 0.6%.**

```{r Customer_churn}
df <- churn_data %>% 
  group_by(churn) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))
str(df)

ggplot(df, aes(x = "", y = perc, fill = churn)) +
  geom_col(color="black") +
  geom_label(aes(label = labels),color = c("black", "black"),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y")+
  scale_fill_brewer(palette="Purples")+
  labs(title="Percentage of customer retained",x="", y="") 
```  

**The bank managed to retain 80% of their customers with the remaining 20% where churned out.**

## Inferential statstics

**Inferential statistics for customer credit scores**

```{r,means and SD for credit score,results='apois'}
mean.credit_score = mean(churn_data$credit_score, na.rm = TRUE); #mean of the customer credit scores
str(mean.credit_score)
sd_credit_score = sd(churn_data$credit_score)
str(sd_credit_score)

```
Mean  and SD for the entire population of credit_score in churn_data data frame mean=`r format(mean.credit_score, digits=5)` and sd=`r format(sd_credit_score, digits=5)`.

Now let us create a smaller sample of the credit score out of the entire population, and call it churn_credit_sample.
```{r churn_credit_sample,results='apois'}
set.seed(321) 
churn_credit_sample = churn_data[ sample(nrow(churn_data),4000), ]
format(mean(churn_credit_sample$credit_score), digits=5)

```

```{r Z interval for customer credit scores,}
loadPkg("BSDA") 

ztestcredit95 = z.test(churn_credit_sample$credit_score, sigma.x = 96.7) # default conf.level = 0.95
ztestcredit95
ztestcredit99 = z.test(churn_credit_sample$credit_score, sigma.x = 96.7, conf.level=0.99 )
ztestcredit99
ztestcredit50 = z.test(churn_credit_sample$credit_score, sigma.x = 96.7, conf.level=0.50 )
ztestcredit50

ztestcredit95$conf.int
ztestcredit99$conf.int
ztestcredit50$conf.int
```

```{r T interval for customer credit scores,}
loadPkg("BSDA") 

ttestcredit95 = t.test(churn_credit_sample$credit_score) # default conf.level = 0.95
ttestcredit95
ttestcredit99 = t.test(churn_credit_sample$credit_score, conf.level=0.99 )
ttestcredit99
ttestcredit50 = t.test(churn_credit_sample$credit_score,  conf.level=0.50 )
ttestcredit50

ttestcredit95$conf.int
ttestcredit99$conf.int
ttestcredit50$conf.int
```

**Inferential statistics for customer account balance.**

```{r,means and SD for  account balance,results='apois'}
mean.balance= mean(churn_data$balance, na.rm = TRUE); #mean of the customer account balance
str(mean.balance)
sd_balance = sd(churn_data$balance)
str(sd_balance)

```

Mean  and SD for the entire population of credit_score in churn_data data frame mean=`r format(mean.balance, digits=5)` and sd=`r format(sd_balance, digits=5)`.

Now let us create a smaller sample of the account balance out of the entire population, and call it churn_balance_sample.

```{r churn_balance_sample,results='apois'}
set.seed(321) 
churn_balance_sample = churn_data[ sample(nrow(churn_data),4000), ]
format(mean(churn_balance_sample$balance), digits=5)

```

```{r Z interval for customer account balance,}
loadPkg("BSDA") 

ztestbalance95 = z.test(churn_balance_sample$balance, sigma.x = 62397.4) # default conf.level = 0.95
ztestbalance95
ztestbalance99 = z.test(churn_balance_sample$balance, sigma.x = 62397.4, conf.level=0.99 )
ztestbalance99
ztestbalance50 = z.test(churn_balance_sample$balance, sigma.x = 62397.4, conf.level=0.50 )
ztestbalance50

ztestbalance95$conf.int
ztestbalance99$conf.int
ztestbalance50$conf.int
```

```{r T interval for customer account balance,}
loadPkg("BSDA") 

ttestbalance95 = t.test(churn_balance_sample$balance) # default conf.level = 0.95
ttestbalance95
ttestbalance99 = t.test(churn_balance_sample$balance, conf.level=0.99 )
ttestbalance99
ttestbalance50 = t.test(churn_balance_sample$balance,  conf.level=0.50 )
ttestbalance50

ttestbalance95$conf.int
ttestbalance99$conf.int
ttestbalance50$conf.int
```


## Churned Customer Analysis

### Mid term Analysis

### SMART Questions:

### Does churn depend on Gender?

**We are doing chi-square test to determine if the churn rate depends on Gender?**

Chi Square Test for independence

H0: Customer’s gender and churn are independent

H1: Customer’s gender and churn are not independent


```{r Chisq test of independence between Gender and customer churn rate}
#Contingency table
contab = table(churn_data$churn, churn_data$gender)
xkabledply(contab, title="Contingency table for Gender(Male, Female) vs Churn (0 or 1)")

#Chisquare Test of independency
chitests = chisq.test(contab)
chitests

chitests$statistic
chitests$parameter
chitests$p.value
```

We have $\chi^2$  value of the test from `chitests$statistic` = `r chitests$statistic`,while extracting the p-value from `chitests$p.value` = `r chitests$p.value` which is less than significance level of 0.05 ,Hence we reject the null hypothesis H0 and we conclude that the churn rate dependent on gender of the customer.

```{r Gender_churns}
ggplot(data=churn_data,aes(x=gender,fill=churn))+
  geom_bar(col="black")+
   scale_fill_manual('Position', values=c('red', 'lightblue'))+
  labs(title = "Gender vs Churn",x="Gender",y="Chrun")+
  theme_minimal()

```   

We observed that majority of the churned customers were female despite the total population of the bank being predominantly male.

### Does a user having a credit card with the bank affect their predisposition to churn? 

Chi Square Test for independence

H0: Having a credit card and churn are independent

H1: Having a credit card and churn are not independent

```{r Chisq test of independence between credit card and customer churn rate}
churn_data[, 10:11][churn_data[, 10:11] == 1] <- 'Churn'
churn_data[, 10:11][churn_data[, 10:11] == 0] <- 'No Churn'

cc_churn <- table(churn_data$churn, churn_data$credit_card)
cc_churn

chitestccchurn <- chisq.test(cc_churn)
chitestccchurn



ggplot(data=churn_data,aes(x=credit_card,fill=churn))+
  geom_bar()+
  scale_fill_manual('Position', values=c('red', 'lightblue')) +
  labs(title = "Customer Churn in Customers With and Without Credit Cards",x="Credit Card",y="Churn")+
  theme_minimal()
```

Credit Card vs Churn
P-value of 0.5 is much higher that the significance level (0.05 for df=1). Thus, we reject the null hypothesis H0, as there is a 0.5 or 50% chance of these results occurring by chance. 

### Does the churn rate depend upon which country the customer belongs to?

Chi Square Test for independence

H0: Country and churn are independent

H1: Country and churn are not independent




```{r Chisq test of independence between country and customer churn rate}
#Contingency table

contable = table(churn_data$country, churn_data$churn)
xkabledply(contable, title="Contingency table for Country (Customer belongs to) vs Churn (0 or 1)")

#Chisquare Test of independency
chitest <- chisq.test(contable)
chitest

chitest$statistic
chitest$parameter
chitest$p.value
### 7290530e7885e0b0a98352236618e34354265333
```

We have the $\chi^2$ value of the test from `chitests$statistic` = `r chitests$statistic`, while the p-value is less than the significance level of 0.05 from `chitests$p.value` = `r chitests$p.value`. As a result, we reject the null hypothesis H0 and find that the churn rate is reliant on the customer's regions.

```{r Country_churns}
ggplot(data=churn_data,aes(x=country,fill=churn))+
  geom_bar()+
  scale_fill_manual('Position', values=c('red', 'lavenderblush2')) +
  labs(title = "Customer Churn in Different Countries",x="Country",y="Churn")+
  theme_minimal()
``` 

We can notice that most of the customers from germany and france are about to churned when compared to customers from spain.

### Does churn depends on different banking product and services provide by the banks?

Chi Square Test for independence

H0: Banking products, services and churn are independent

H1: Banking products, services and churn are not independent


```{r Chisq test of independence between Banking product and services and customer churn rate}
#Contingency table
contab_productnumber = table(churn_data$churn, churn_data$products_number)
xkabledply(contab_productnumber, title="Contingency table for Banking Services vs Churn (0 or 1)")

#Chisquare Test of independency
chitests = chisq.test(contab_productnumber)
chitests

chitests$statistic
chitests$parameter
```

We have the $\chi^2$ value of the test from `chitests$statistic` = `r chitests$statistic`, while the p-value is less than the significance level of 0.05 from `chitests$p.value` = `r chitests$p.value`. As a result, we reject the null hypothesis H0 and find that the churn rate is reliant on the customer's active status.

```{r Bank_Services_churns}
ggplot(data=churn_data,aes(x=products_number,fill=churn))+
  geom_bar()+
  scale_fill_manual('Position', values=c('red', 'lavender')) +
  labs(title = "Banking Product and Services vs Churn",x="Banking Product and Services",y="Chrun")+
  theme_minimal()
```  

We can notice that most of the customers with lower and higher products i.e) 1, 3 and 4 are about to churned when compared to customers with two products.

### Does churn depend on status of active account users?

Chi Square Test for independence

H0: Customer’s active status and churn are independent

H1: Customer’s active status  and churn are not independent

```{r Chisq test of independence between Active members and customer churn rate}
#Contingency table
contab_activemember = table(churn_data$churn, churn_data$active_member)
xkabledply(contab_activemember, title="Contingency table for Account Status(Active, In Active) vs Churn (0 or 1)")

#Chisquare Test of independency
chitests = chisq.test(contab_activemember)
chitests

chitests$statistic
chitests$parameter
chitests$p.value
```

We have the $\chi^2$ value of the test from `chitests$statistic` = `r chitests$statistic`, while the p-value is less than the significance level of 0.05 from `chitests$p.value` = `r chitests$p.value`. As a result, we reject the null hypothesis H0 and find that the churn rate is reliant on the customer's banking services.

```{r Active_members_churns}
ggplot(data=churn_data,aes(x=active_member,fill=churn))+
  geom_bar()+
  scale_fill_manual('Position', values=c('red', 'pink1')) +
  labs(title = "Active Members vs Churn",x="Active Status of Account Holders",y="Churn")+
  theme_minimal()
```  

Here, we can interpret that most of the Inactive customers are about to churned.

### Does a higher/lower account balance of the customer affect churn? 

Correlation test between account balance and churn

```{r Correlation between balance and customer churn, include=FALSE, echo=FALSE}
library(corrplot)
churn_data_to_numeric <- churn_data

churn_data$churn[churn_data$churn== "Churned"]<-1
churn_data$churn[churn_data$churn== "Retained"]<-0


churn_data_to_numeric$churn <- as.numeric(churn_data_to_numeric$churn) 

churn_data_numeric <- select_if(churn_data_to_numeric, is.numeric)

cors_bal <- cor(churn_data_numeric)
corrplot(cors_bal, method = 'square')

cor_balance <- cor(churn_data_numeric$balance, churn_data_numeric$churn)
cor_balance
```

```{r Correlation between balances and customer churn, include=TRUE, echo=TRUE}
library(corrplot)
churn_data_to_numeric <- churn_data
churn_data_to_numeric$churn <- as.numeric(churn_data_to_numeric$churn) 

churn_data_numeric <- select_if(churn_data_to_numeric, is.numeric)

cors_bal <- cor(churn_data_numeric)
corrplot(cors_bal, method = 'number')
cor_balance<-cor.test(churn_data$balance,as.numeric(churn_data$churn), method="pearson")
cor_balance

```  

Customer churn is weakly correlated with account balance with a score of 0.12
For further analysis we can check the correlation score by combining multiple variables 


```{r}
loadPkg("BSDA") 

ttestbalances95 = t.test(churn_balance_sample$credit_score) # default conf.level = 0.95
ttestbalances95
ttestbalances99 = t.test(churn_balance_sample$credit_score, conf.level=0.99 )
ttestbalances99
ttestbalances50 = t.test(churn_balance_sample$credit_score,  conf.level=0.50 )
ttestbalances50

ttestbalances95$conf.int
ttestbalances99$conf.int
ttestbalances50$conf.int
```

### Does a higher/lower credit score of the customer affect churn?

Correlation test between credit score and churn

```{r Correlation between credit score and customer churn, include=TRUE, echo=TRUE}
library(corrplot)
churn_data_to_numeric <- churn_data

churn_data$churn[churn_data$churn== "Churned"]<-1
churn_data$churn[churn_data$churn== "Retained"]<-0


churn_data_to_numeric$churn <- as.numeric(churn_data_to_numeric$churn) 
str(churn_data_to_numeric)
churn_data_numeric <- select_if(churn_data_to_numeric, is.numeric)
str(churn_data_numeric)
cors_credit <- cor(churn_data_numeric)
corrplot(cors_credit, method = 'number')

cor_scores <- cor(churn_data_numeric$credit_score, churn_data_numeric$churn)
cor_scores

```


**The credit score is weakly correlated with customer churn with a score of -0.0271**

### Does churn depend on age of the customers?
```{r Correlation between age and customer churn, include=TRUE, echo=TRUE}
library(corrplot)
churn_data_to_numeric <- churn_data

churn_data$churn[churn_data$churn== "Churned"]<-1
churn_data$churn[churn_data$churn== "Retained"]<-0


churn_data_to_numeric$churn <- as.numeric(churn_data_to_numeric$churn) 
str(churn_data_to_numeric)
churn_data_numeric <- select_if(churn_data_to_numeric, is.numeric)
str(churn_data_numeric)
cors_age_group <- cor(churn_data_numeric)
corrplot(cors_age_group, method = 'number')

cor_age <- cor(churn_data_numeric$age, churn_data_numeric$churn)
cor_age
```
  
```{r}
loadPkg("BSDA") 

ttestage95 = t.test(churn_balance_sample$age) # default conf.level = 0.95
ttestage95
ttestage99 = t.test(churn_balance_sample$age, conf.level=0.99 )
ttestage99
ttestage50 = t.test(churn_balance_sample$age,  conf.level=0.50 )
ttestage50

ttestage95$conf.int
ttestage99$conf.int
ttestage50$conf.int
```

```{r}

ggplot(churn_data, aes(x=age, fill=churn)) +
  geom_histogram( color='#e9ecef', alpha=0.5, position='identity') + scale_fill_manual('Position', values=c('red', 'pink1')) +
  labs(title = "Age Distribution across Churned and No Churn Account Holders",x="Age of Account Holders",y="Frequency")

#Contingency table
churn_age = table(churn_data$churn, churn_data$age)
churn_age

#Chisquare Test of independence
chichurn_age = chisq.test(churn_age)
chichurn_age

chichurn_age$statistic
chichurn_age$parameter
chichurn_age$p.value

```


### Final Project Analysis

***We observed that correlation of age variable is very high,so we decided to combine different columns and find if the correlation score improves or not**

```{r}
corre<-cor(churn_data_numeric)
corre
corrplot(corre, method = "number")
```

Churn & Age are about to moderately correlated (Positive) - 0.29.

```{r Correlation combining multiple variables, results='markup', include=TRUE}
corr <- cor(churn_data_numeric$age, churn_data_numeric$churn)
corr
corr1 <- cor(churn_data_numeric$age + churn_data_numeric$tenure, churn_data_numeric$churn)
corr1
corr2 <- cor(churn_data_numeric$age + churn_data_numeric$products_number, churn_data_numeric$churn)
corr2
corr3 <- cor(churn_data_numeric$age + churn_data_numeric$balance, churn_data_numeric$churn)
corr3
corr4 <- cor(churn_data_numeric$age + churn_data_numeric$tenure + churn_data_numeric$products_number, churn_data_numeric$churn)
corr4
corr5 <- cor(churn_data_numeric$age + churn_data_numeric$products_number + churn_data_numeric$balance, churn_data_numeric$churn)
corr5

```
The correlation score is the highest for the age variable. After combining different columns and finding the correlation score we see that it starts decreasing from 0.285

***Since data set is unbalanced we decided to use oversampling technique to balance the dataset using library called ROSE***

The data set was unbalance majorly because the variable country had most of the data which belongs to France,Hence we need to do perform data balancing.

```{r count of each country before oversampling}
table(churn_data$churn)
table(churn_data$country)
```

```{r, Balancing data using oversampling, results='markup', include=TRUE}
library(ROSE)

sum(churn_data$country == 'France')
sum(churn_data$country == 'Germany')
sum(churn_data$country == 'Spain')
ind <- sample(2, nrow(churn_data), replace = TRUE, prob = c(0.7, 0.3))
train <- churn_data[ind==1,]
test <- churn_data[ind==2,]
```

```{r count of each country after oversampling}
data_balanced_over <- ovun.sample(churn ~ ., data = train, method = "over",N = 10000,seed=123)$data
table(data_balanced_over$churn)
table(data_balanced_over$country)
```

```{r}
ggplot(data=data_balanced_over,aes(x=country,fill=country))+
  geom_bar(col="black")+
  scale_fill_brewer(palette="Reds") +
  labs(title = "Bank Customer vs Country",x="Country",y="Customer Count")+
  theme_minimal()
```

### Model Building and trainig

**Let us see the probability of customer are churned/not churned.**

```{r}
prop.table( table(churn_data$churn) )
```


we observe that his probability table tells you that around 20 percent of the customers are churned. If those customers are all the ones that are having good relationship with the bank, then the bank is going to face severe strike on its business rate. Let’s use the different modelling methods to train our dataset to see if we can find out what’s causing customers to close the account with the particular bank.

```{r, Scaling data}
str(churn_data)
churn_data_pc<-data_balanced_over
churn_data_pc$country[churn_data$country == 'France'] <- 0
churn_data_pc$country[churn_data$country == 'Germany'] <- 1
churn_data_pc$country[churn_data$country == 'Spain'] <- 2

churn_data_pc$gender[churn_data$gender == 'Male'] <- 0
churn_data_pc$gender[churn_data$gender == 'Female'] <- 1

churn_data_pc$credit_card[churn_data$credit_card == 'No-Credit Card'] <- 0
churn_data_pc$credit_card[churn_data$credit_card == 'Credit Card'] <- 1

churn_data_pc$active_member[churn_data$active_member == 'In Active'] <- 0
churn_data_pc$active_member[churn_data$active_member == 'Active'] <- 1

churn_data_pc$active_member = as.numeric(churn_data_pc$active_member)
churn_data_pc$credit_card = as.numeric(churn_data_pc$credit_card)
churn_data_pc$country = as.numeric(churn_data_pc$country)
churn_data_pc$credit_score = as.numeric(churn_data_pc$credit_score)
churn_data_pc$gender = as.numeric(churn_data_pc$gender)
churn_data_pc$churn = as.numeric(churn_data_pc$churn)

churn_data_pc_target = churn_data_pc #includes 'churn'
```

```{r}
churn_data_logit <- churn_data_pc_target
str(churn_data_logit)
```

```{r}                                        
churn_data_pc <- churn_data_logit[-c(11)] #No churn
str(churn_data_pc)
```

```{r}
str(churn_data_pc_target)
str(churn_data_pc)
churn_data_scale <- data.frame(scale(churn_data_pc))
xkabledply(cor(churn_data_scale))
xkabledply(cov(churn_data_scale))

```
### SMART QUESTIONS

### Does the combination of customer demographic variables country, age group, gender influence the churn rate?

```{r LMmodel1,results='markup'}
churnLM <- lm(churn~country+age+gender,data=churn_data_pc_target)
summary(churnLM)
xkabledply(churnLM, title = "churn Model summary")

confint(churnLM)
```

### Whether the customers using fewer products with higher salary,account balance and status of the account affect the churn rate?

```{r LMmodel2,results='markup'}
churnLM2 <- lm(churn~products_number+estimated_salary+balance+active_member,data=churn_data_pc_target)
summary(churnLM2)
xkabledply(churnLM2, title = "churn Model summary")


confint(churnLM2)
```

###	What are the principal components for predicting churn rate?

```{r, Proportion of variance}
pr.out=prcomp(churn_data_pc, scale =TRUE)
print("Case: z-score/scaled")
summary(pr.out)
pr.out$rotation
```

```{r}
nrow(train)
prop.table( table(train$churn) )
```

```{r, Biplot}
biplot(pr.out, scale = 0)
biplot(pr.out,2:3, scale =0)
biplot(pr.out,3:4, scale =0)
```


```{r}
pr.var <- (pr.out$sdev^2)
pve <- pr.var/sum(pr.var)
plot(cumsum(pve), xlab="Principal Component (standardized)", ylab ="Cumulative Proportion of Variance Explained",ylim=c(0,1),type="b")
```

### Which model can give the best results for adjusted R square and ROC value with lower AIC and BIC?

**Feature selection using Exhaustive search**

```{r}
loadPkg("leaps")

reg.churn <- regsubsets(churn~., data = churn_data_pc_target, nvmax =9, nbest = 1, method = "exhaustive")  
plot(reg.churn , scale = "adjr2", main = "Adjusted R^2")
plot(reg.churn , scale = "r2", main = "R^2")
plot(reg.churn , scale = "bic", main = "BIC")
plot(reg.churn , scale = "Cp", main = "Cp")
summary(reg.churn )
```

Using exhaustive search age,balance were selected  which is a two variable model with adjusted R^2 value 0.14.

The  best model selected using BIC is credit_score,age,balance which is a 3 variable model with BIC value -1500.

The best model selected using Cp is credit_score,age,tenure,balance,estimated_salary which is a 5 variable model. with Cp value 7.


```{r}
loadPkg("car")

# Adjusted R2
car::subsets(reg.churn, statistic="adjr2", legend = FALSE, min.size = 2, main = "Adjusted R^2")

subsets(reg.churn, statistic="cp", legend = FALSE, min.size = 4, main = "Mallow Cp")
abline(a = 1, b = 1, lty = 3)  
```

The Mallow Cp plot selected tw0 best models.

1)credit_score,country,gender,age,tenure,balance,active_member,estimated salary 8 variable model
2)credit_score,country,gender,age,tenure,balance,estimated_salary 7 variable model.


**Feature selection using forward search**

```{r}
reg.churnforward <- regsubsets(churn~., data = churn_data_pc_target, nvmax =9, nbest = 1, method = "forward")
plot(reg.churnforward , scale = "adjr2", main = "Adjusted R^2")
plot(reg.churnforward , scale = "bic", main = "BIC")
plot(reg.churnforward , scale = "Cp", main = "Cp")
summary(reg.churnforward)
```

The best model selected using forward search is age,balance which is a 2 variable model with adjusted R^2 value of 0.14.

The  best model selected using BIC is credit_card,age,balance which is a 3 variable model with value -1500 Cp.

The best model selected using Cp is credit_score,age,tenure,balance,estimated_salary which is a 5 variable model with Cp value 7.

**Feature selection using backward search**

```{r}
reg.churnback <- regsubsets(churn~., data = churn_data_pc_target, method = "backward", nvmax = 9, nbest = 1)
plot(reg.churnback, scale = "adjr2", main = "Adjusted R^2")
plot(reg.churnback, scale = "bic", main = "BIC")
plot(reg.churnback, scale = "Cp", main = "Cp")
summary(reg.churnback)
```

The best model selected using backward search is age,balance which is a 2 variable model with adjusted R^2 value of 0.14.

The  best model selected using BIC is credit_card,age,balance which is a 3 variable model with value -1500 Cp.

The best model selected using Cp is credit_score,age,tenure,balance,estimated_salary which is a 5 variable model with Cp value 7.

**Feature selection using Sequential Replacement**

```{r}
reg.churnseqrep <- regsubsets(churn~., data =churn_data_pc_target, nvmax = 9, nbest = 1 , method = "seqrep")
plot(reg.churnseqrep, scale = "adjr2", main = "Adjusted R^2")
plot(reg.churnseqrep, scale = "bic", main = "BIC")
plot(reg.churnseqrep, scale = "Cp", main = "Cp")
```

The best model selected using sequential search is age,balance which is a 2 variable model with adjusted R^2 value of 0.14.

The  best model selected using BIC is credit_card,age,balance,estimated_salary which is a 4 variable model with value -1500 Cp.

The best model selected using Cp is credit_score,age,tenure,balance,estimated_salary which is a 5 variable model with Cp value 7.


### Logistic Regression model

### What factors affect the customer churn rate significantly?

```{r, Logistic Regression,results='markup'}

churn_logit <- glm(churn ~., data = churn_data_logit, family = 'binomial')

summary(churn_logit)
```

```{r ConfInt_FS, results='markup', collapse=F}
xkabledply( confint(churn_logit), title = "CIs using profiled log-likelihood" )
xkabledply( confint.default(churn_logit), title = "CIs using standard errors" )
```

```{r confusionMatrix_FS, results='markup'}
loadPkg("regclass")
xkabledply( confusion_matrix(churn_logit), title = "Confusion matrix from Logit Model" )
```

```{r roc_auc_FS, results='markup'}
loadPkg("pROC") 
prob=predict(churn_logit, type = "response" )
churn_data_logit$prob=prob
h <- roc(churn~prob, data = churn_data_logit)
auc(h) 
plot(h)
```

Based on the models obtained from feature selection we are using logistic regression to check the churn rate of the customers.

```{r Model1,results='markup'}
churnglm1 <- glm(churn ~  age + balance, data = churn_data_logit)
summary(churnglm1)
xkabledply( confusion_matrix(churnglm1), title = "Confusion matrix from Logit Model-1" )
prob1=predict(churnglm1, type = "response" )
churn_data_logit$prob=prob1
h1 <- roc(churn~prob, data = churn_data_logit)
auc(h1) 
plot(h1)
```

```{r Model2,results='markup'}
churnglm2 <- glm(churn ~  credit_card+age + balance, data = churn_data_logit)
summary(churnglm2)
xkabledply( confusion_matrix(churnglm2), title = "Confusion matrix from Logit Model-2" )
prob2=predict(churnglm2, type = "response" )
churn_data_logit$prob=prob2
h2 <- roc(churn~prob, data = churn_data_logit)
auc(h2) 
plot(h2)
```

```{r Model3,results='markup'}
churnglm3 <- glm(churn ~ credit_score+age+tenure+ balance + estimated_salary, data = churn_data_logit)
summary(churnglm3)
xkabledply( confusion_matrix(churnglm3), title = "Confusion matrix from Logit Model-3" )
prob3=predict(churnglm3, type = "response" )
churn_data_logit$prob=prob3
h3 <- roc(churn~prob, data = churn_data_logit)
auc(h3) 
plot(h3)
```


```{r Model4,results='markup'}
churnglm4 <- glm(churn ~ credit_score+country+gender+age+tenure+balance+estimated_salary, data = churn_data_logit)
summary(churnglm4)
xkabledply( confusion_matrix(churnglm4), title = "Confusion matrix from Logit Model-4" )
prob4=predict(churnglm4, type = "response" )
churn_data_logit$prob=prob4
h4 <- roc(churn~prob, data = churn_data_logit)
auc(h4) 
plot(h4)
```

```{r Model5,results='markup'}
churnglm5 <- glm(churn ~ credit_score + country+ gender+age+tenure+balance+active_member+estimated_salary
, data = churn_data_logit)
summary(churnglm5)
xkabledply( confusion_matrix(churnglm5), title = "Confusion matrix from Logit Model-5" )
prob5=predict(churnglm5, type = "response" )
churn_data_logit$prob=prob5
h5 <- roc(churn~prob, data = churn_data_logit)
auc(h5) 
plot(h5)
```










